---
# bibliography: NEriskpolicy.bib
csl: plos.csl
fontsize: 10pt
geometry: left=2cm, right=2cm, top=2cm, bottom=3cm, footskip = .5cm
link-citations: yes
output:
  bookdown::pdf_document2:
    includes: 
       in_header: latex/header.tex
    keep_tex: true
  bookdown::html_document2: 
    toc: true
    toc_float: true
    code_fold: hide
  bookdown::word_document2:
    toc: true
subparagraph: yes
urlcolor: blue
---

```{r setup, include=FALSE}

# library(tint)
# # invalidate cache when the package version changes
# knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
# options(htmltools.dir.version = FALSE)

#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      #dev = "cairo_pdf",
                      warning = FALSE,
                      fig.width = 4,
                      fig.asp = 0.45,
                      fig.align = 'center'
                      ) #allows for inserting R code into captions

#Plotting and data libraries
#remotes::install_github("noaa-edab/ecodata@0.1.0") #change to 2020 ecodata version for release
library(tidyverse)
library(tidyr)
library(ecodata)
library(here)
library(kableExtra)
library(patchwork)

```


# New England Risk Policy



## Potential indicators for Climate and Ecosystem Considerations

**American plaice indicators**

```{r plaice_eco, echo=FALSE}

ecodata::plot_bottom_temp(report = "NewEngland")

```

**Atlantic cod indicators**

```{r cod_eco, echo=FALSE}
source(here::here('04_scripts/cod_ecosystem_indicators.R'))

```

**Atlantic herring indicators**

```{r herring_eco, echo=FALSE}

source(here::here('04_scripts/herring_ecosystem_indicators.R'))

```

\*\* Georges Bank Yellowtail flounder indicators\*\*

```{r gb_yt_eco, echo=FALSE}

ecodata::plot_bottom_temp(report = "NewEngland")

```

\*\* Southern New England Yellowtail flounder indicators\*\*

```{r sne_yt_eco, echo=FALSE}

ecodata::plot_cold_pool(report = "MidAtlantic", varName = 'persistence')

```

## Potential Indicators for Commercial Fishery Characterization

**Commercial fishery characterization indicators**

```{r comm, fig.cap='Effective Shannon index of NE commercial fisheries' ,echo=FALSE}

# call in commercial fishery diversity data
data <- ecodata::commercial_div

# plot fleet diversity in revenue for New England
data |> 
  dplyr::filter(Var == "Fleet diversity in revenue") |>
  dplyr::filter(EPU == "NE") |>
  #plot Value by Time
  ggplot2::ggplot(ggplot2::aes(x = Time, y = Value)) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::ylab("Effective Shannon index")
```

## Potential Indicators for Recreational Fishery Characterization

**Recreational fishery characterization indicators**

```{r rec, echo=FALSE}

ecodata::plot_recdat(report = "NewEngland", varName = 'effortdiversity')

```
