---
# bibliography: NEriskpolicy.bib
csl: plos.csl
fontsize: 10pt
geometry: left=2cm, right=2cm, top=2cm, bottom=3cm, footskip = .5cm
link-citations: yes
output:
  bookdown::pdf_document2:
    includes: 
       in_header: latex/header.tex
    keep_tex: true
  bookdown::html_document2: 
    toc: true
    toc_float: true
    code_fold: hide
  bookdown::word_document2:
    toc: true
subparagraph: yes
urlcolor: blue
---

```{r setup, include=FALSE}

# library(tint)
# # invalidate cache when the package version changes
# knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
# options(htmltools.dir.version = FALSE)

#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      #dev = "cairo_pdf",
                      warning = FALSE,
                      fig.width = 4,
                      fig.asp = 0.45,
                      fig.align = 'center'
                      ) #allows for inserting R code into captions

#Plotting and data libraries
#remotes::install_github("noaa-edab/ecodata@0.1.0") #change to 2020 ecodata version for release
library(tidyverse)
library(tidyr)
library(ecodata)
library(here)
library(kableExtra)
library(patchwork)

```

# New England Risk Policy

The Risk Policy Statement and Concept (2025) outlines how the Council intends to operationalize environmental and socioeconomic considerations in management. The Council also highlights how they anticipate organizations, including the Northeast Fisheries Science Center (NEFSC), contributing to this effort. The NEFSC is expected to contribute through continued updates of the State of the Ecosystem report, and further development of risk indicators. Risk Policy Factors developed by the Council consider 1. Stock Status and Uncertainty, 2. Climate and Ecosystem Considerations, and 3. Economic and Community Importance. Factors related to Stock Status and Uncertainty are derived from stock assessments. Here we demonstrate how indicators from the State of the Ecosystem report and indicators developed for research track assessments could inform Climate and Ecosystem Considerations and Economic and Community Importance.

## Potential indicators for Climate and Ecosystem Considerations

In developing indicators for Climate and Ecosystem Considerations, we prioritized indicators with a statistically significant, mechanistic correlation with the life history they were designed to assess. We prioritized indicators which have not been included in the stock assessment to date to avoid redundant consideration of the indicator in management advice. We prioritized indicators from Term of Reference 1 (ToR1) from research track stock assessments as they have been collaboratively developed with stock assessment working groups and peer reviewed.

**American plaice indicators**

A research track assessment of American plaice was completed in 2022. TOR1 of the assessment identified multiple environmental indicators that are correlated with life history of American plaice. The most appropriate indicator for application in the Risk Policy is Annual mean bottom temperature anomaly. This indicator was significantly correlated with fall and spring mean latitude of American plaice in the NEFSC trawl survey, suggesting the species is distributed further North and East in warmer years compared to cooler years. Bottom temperature anomaly is regularly included in the State of the Ecosystem report. Details on this indicator are available in the SOE technical document.

```{r plaice_eco, echo=FALSE}

ecodata::plot_bottom_temp(report = "NewEngland")

```

**Atlantic cod indicators**

```{r cod_eco, echo=FALSE}
source(here::here('04_scripts/cod_ecosystem_indicators.R'))

```

**Atlantic herring indicators**

```{r herring_eco, echo=FALSE}

source(here::here('04_scripts/herring_ecosystem_indicators.R'))

```

\*\* Georges Bank Yellowtail flounder indicators\*\*

```{r gb_yt_eco, echo=FALSE}

ecodata::plot_bottom_temp(report = "NewEngland")

```

\*\* Southern New England Yellowtail flounder indicators\*\*

```{r sne_yt_eco, echo=FALSE}

ecodata::plot_cold_pool(report = "MidAtlantic", varName = 'persistence')

```

## Potential Indicators for Commercial Fishery Characterization

**Commercial fishery characterization indicators**

```{r comm, fig.cap='Effective Shannon index of NE commercial fisheries' ,echo=FALSE}

# call in figures from 05_images
source(here::here('04_scripts/comm_indicators.R'))

print(trend_summary)
```

## Potential Indicators for Recreational Fishery Characterization

**Recreational fishery characterization indicators**

```{r rec, echo=FALSE}

ecodata::plot_recdat(report = "NewEngland", varName = 'effortdiversity')

```
